---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ohseer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of ohseer is to provide an R interface to the Mistral AI OCR (Optical Character Recognition) API. This package allows you to easily extract text from documents using Mistral's powerful OCR capabilities directly from your R environment.

## Part of the EcoExtract Suite


`OhSeeR` is the foundational first step in the **EcoExtract Suite**, a collection of R packages designed for extracting and structuring ecological data from academic literature. This suite facilitates a modular workflow from raw documents to validated, analysis-ready data.

The overall workflow is visualized below:

```{mermaid}
graph TD
    subgraph Input [ ]
        A[(Source PDF Documents)]
    end
    subgraph Processing_Pipeline [ ]
        A -- Raw PDF Data --> B(OhSeeR):::package;
        B -- PDF Text --> C(sanitizeR):::package;
        C -- Cleaned Text --> E{LLM API Call};
        D(whispeR):::package -- Formatted Prompts --> E;
        E -- Raw LLM Response --> F(structuR):::package;
        F -- Structured Data --> H(auditR):::package;
    end
    subgraph Validation_Source [ ]
        C --> H;
    end
    subgraph Output [ ]
        H -- Validated Data --> I((Structured Dataset));
    end
    %% Styling classes for package nodes
    classDef package fill:#D6EAF8,stroke:#5DADE2,stroke-width:2px;
    %% Remove subgraph backgrounds and borders
    classDef subgraphStyle fill:transparent,stroke:transparent;
    class Input,Processing_Pipeline,Validation_Source,Output subgraphStyle;
```

## Features

- Upload files to Mistral AI
- Process documents via URL or uploaded files
- Extract text and images from PDFs
- Preview OCR results as HTML with embedded images
- Embed base64 images in markdown for Shiny apps
- Simple, consistent interface
- No heavy image processing dependencies (no magick required)

## Installation

You can install the development version of ohseer from [GitHub](https://github.com/) with:

```r
# Option 1: Using pak (recommended)
# install.packages("pak")
pak::pak("n8layman/ohseer")

# Option 2: Using devtools
# install.packages("devtools")
devtools::install_github("n8layman/ohseer")

# Option 3: Using remotes
# install.packages("remotes")
remotes::install_github("n8layman/ohseer")
```

## Authentication

To use this package, you’ll need a Mistral AI API key. To receive an api key you will need to:

1. Visit [mistral.ai](https://mistral.ai/)
2. Click 'Try the API' from the top menu bar
3. Sign in using Google, Apple, or Microsoft accounts or register
4. Create and name a workspace or organization
5. Click 'Subscription' and then 'Compare plans' from the left task bar.
6. Chose 'Experiment for Free' to try out the service and subscribe. This will require entering your phone number for validation
7. Click on 'API keys' and 'Create New Key`
8. Copy down your API key and set it as an environment variable in R

```r
Sys.setenv(MISTRAL_API_KEY = "your-api-key-here")
```

For persistent authentication, add this to your `.Renviron` file:

```r
MISTRAL_API_KEY=your-api-key-here
```

⚠️ **Security Warning**: Never commit `.env` or `.Renviron` files containing API keys to version control. Consider using encrypted environment files or a secure credential management system for production environments.

## Examples

### Basic OCR Processing

The simplest way to process a document is to provide a URL:

```{r eval=FALSE}
library(ohseer)

# OCR processing of a PDF from a URL
result <- mistral_ocr("https://arxiv.org/pdf/2201.04234.pdf")
```

### Processing Local Files

You can also process local documents:

```{r eval=FALSE}
# Process a local PDF file
result <- mistral_ocr("path/to/document.pdf")

# Save the OCR results to a file
result <- mistral_ocr("path/to/document.pdf", output_file = "ocr_result.json")
```

### Working with File IDs

If you already have a file ID from a previous upload:

```{r eval=FALSE}
# Retrieve a file using its ID
file_content <- mistral_ocr("00edaf84-95b0-45db-8f83-f71138491f23")
```

### Previewing OCR Results

You can preview OCR results with embedded images:

```{r eval=FALSE}
# Process a document
result <- mistral_ocr("path/to/document.pdf")

# Preview a page as HTML with embedded images
mistral_preview_html(result, page_num = 1)

# For Shiny apps: embed images in markdown
processed_markdown <- mistral_embed_images(
  markdown_text = result$pages[[1]]$markdown,
  mistral_response = result,
  page_num = 1
)

# Then render with your preferred markdown renderer
html_output <- commonmark::markdown_html(processed_markdown)
```

### Direct API Access

For more control, you can use the lower-level functions:
```{r eval=FALSE}
# Upload a file
upload_result <- mistral_ocr_upload_file("path/to/document.pdf")
file_id <- upload_result$id

# Process a URL directly
url_result <- mistral_ocr_url("https://arxiv.org/pdf/2201.04234.pdf")

# Retrieve a file by ID
file_content <- mistral_ocr_retrieve_file(file_id, output_path = "retrieved_document.pdf")
```

## API Functions

### Core OCR Functions

- `mistral_ocr()`: Main function that auto-detects input type (URL, file, or file ID)
- `mistral_ocr_url()`: Process a document at a URL
- `mistral_ocr_upload_file()`: Upload a file to Mistral AI
- `mistral_ocr_retrieve_file()`: Retrieve a file from Mistral AI using its ID

### Preview and Display Functions

- `mistral_preview_html()`: Generate a complete HTML preview with embedded images
- `mistral_embed_images()`: Embed base64 images into markdown text (for Shiny apps)
- `mistral_preview_page()`: Simple markdown-to-HTML preview (without images)
- `mistral_preview_image()`: Display individual images from OCR results

## Notes

- This package is experimental and the API may change
- Large files may take some time to process
- Check Mistral AI documentation for the latest API information

## License

This package is licensed under the MIT License.
